# Mutually Exclusive Spectral Biclustering (MESBC) 

We simulate an example of gene expression count data with five biclusters with the following size: 20×20, 50×20, 50×20, 30×40, and 100×50. It is assumed that the non-bicluster elements (i.e., the background noise) are randomly generated from a negative binomial distribution with a mean of mu0=0.1, and the elements within biclusters are sampled from a negative binomial distribution with a mean of mu=1. After generating data, shuffle it and record and save the real labels.


```{r}
library(pheatmap)
library(scran)
library(igraph)
library(cluster)
library(mclust)
library(clue)
library(ggplot2)

##### Function of generating negative binomial distribution simulation data
gendata_NBP=function(mu,mu0,size,rownum,colnum){
  K=length(rownum)
  N=sum(rownum)
  P=sum(colnum)
  n=rownum
  p=colnum
  
  noise_data=matrix(rnbinom(N*P, mu = mu0, size = size),nrow=N,ncol=P) 
  
  for(i in 1:K){
    
    x=matrix(rnbinom(n[i]*p[i], mu = mu[i], size = size),nrow=n[i],ncol=p[i])
    
    if(i==1) noise_data[1:n[i],1:p[i]]=x
    if(i>1) noise_data[(sum(n[1:(i-1)])+1):sum(n[1:i]),(sum(p[1:(i-1)])+1):sum(p[1:i])]=x
  }
  
  return(noise_data)
  
}

rbicnum=c(20,50,50,30,100)
cbicnum=c(20,20,20,40,50)
mu=rep(1,5)
mu0=0.1
size=1
N=100
ngene=sum(rbicnum)
ncond=sum(cbicnum)
rowend=colend=0
for(i in 1:length(rbicnum)){
  rowend[i+1]=sum(rbicnum[1:i])
  colend[i+1]=sum(cbicnum[1:i])
}
truebic2=list()
for(i in 1:(length(rowend)-1)){
  truebic2$bicr[[paste0("bic",i)]]=(rowend[i]+1):rowend[i+1]
  truebic2$bicc[[paste0("bic",i)]]=ngene+(colend[i]+1):colend[i+1]
}
data=gendata_NBP(mu,mu0,size,rbicnum,cbicnum)
m=nrow(data)
p=ncol(data)
n=m+p
indr=sample(1:m,m,replace = F)
indc=sample(1:p,p,replace = F)
truebic=truebic2
for(i in 1:(length(rowend)-1)){
    for(j in 1:length(truebic2[["bicr"]][[paste0("bic",i)]]))
      truebic[["bicr"]][[paste0("bic",i)]][j]=which(indr==truebic2[["bicr"]][[paste0("bic",i)]][j])
    
    for(j in 1:length(truebic2[["bicc"]][[paste0("bic",i)]]))
      truebic[["bicc"]][[paste0("bic",i)]][j]=which((indc+ngene)==truebic2[["bicc"]][[paste0("bic",i)]][j])+ngene
    
  }
data2=data[indr,indc]

```


We use the MESBC method to perform bicluster analysis on the shuffled matrix. Traversing the results of clustering K from 2 to 10, select the clustering result corresponding to the maximum modularity value.

```{r}
MESBC <- function(data, K){
  m = nrow(data)
  p = ncol(data)
  n = m+p
  w = as.matrix(data+abs(min(0,range(data)[1])))
  d1 = apply(w,1,sum)
  d2 = apply(w,2,sum)
  D1 = diag(1/d1^0.5)
  D2 = diag(1/d2^0.5)
  w.standard = D1%*%w%*%D2
  mysvd = svd(w.standard)
  
  clust <- matrix(nrow=n,ncol=length(K))
  colnames(clust)=K
  for(k in K){
    if (k <= p & k <= m){
      u = mysvd$u[,1:k]; v = mysvd$v[,1:k]; lambda = mysvd$d[1:k]
    }
    if (k > p | k > m)
    { u = mysvd$u ; v = mysvd$v }
    u = as.matrix(u); v = as.matrix(v)
    Y = rbind(D1%*%u,D2%*%v)%*%diag(lambda)
    
    clus.out = kmeans(Y,centers=k,iter.max=100,nstart=100)
    clust[,as.character(k)] = clus.out$cluster
  }
  
  if(length(K)==1) clust=clust[,1]
  return(list(clus=clust, Y=rbind(D1%*%mysvd$u,D2%*%mysvd$v)%*%diag(mysvd$d)))
}

K=2:10
bcout=MESBC(data2,K)
g = buildSNNGraph(t(bcout[["Y"]]))
distY=dist(bcout[["Y"]])
  
grouptmp=numeric()
for(i in 1:(length(rowend)-1))
    grouptmp[c(truebic[["bicr"]][[paste0("bic",i)]],truebic[["bicc"]][[paste0("bic",i)]])]=i
mod=numeric()
clus=list()
  
for(k in K){
    clus$MESBC$row[[as.character(k)]]=bcout[["clus"]][1:m,as.character(k)]
    clus$MESBC$col[[as.character(k)]]=bcout[["clus"]][(m+1):n,as.character(k)]
    mod[k]=modularity(g,c(clus$MESBC$row[[as.character(k)]],clus$MESBC$col[[as.character(k)]]))
}
kbest=as.character(which(mod==max(mod,na.rm = T))[1])
```


Now let's take a look at the heatmaps of the raw data, shuffled data, and data rearranged by MESBC.

```{r}
# true block
pheatmap::pheatmap(data,show_colnames =F,show_rownames = F,legend =F,
                   cluster_rows = F,cluster_cols = F,
                   annotation_colors =F,annotation_legend = F)


# disturbed data
pheatmap::pheatmap(data2,show_colnames =F,show_rownames = F,legend =F,
                   cluster_rows = F,cluster_cols = F,
                   annotation_colors =F,annotation_legend = F)

# data rearranged by MESBC
pheatmap::pheatmap(data2[order(clus[["MESBC"]][["row"]][[kbest]]),order(clus[["MESBC"]][["col"]][[kbest]])],
                   show_colnames =F,show_rownames = F,legend =F,
                   cluster_rows = F,cluster_cols = F,
                   annotation_colors =F,annotation_legend = F)


```

It can be seen that MESBC can correctly identify the five designated blocks.


We can also analyze multiple metrics of clustering performance, which requires multiple repeated experiments. The corresponding code is in "code for MESBC analysis.R".
